Process-based phenotypes: an application to the Cellular Phenotype Ontology

### 2011-11-23 paper/drinking.tex
added GO blurb, including regulation stuff

\subsection{Gene Ontology}
quick intro...

>>>>

\subsection{Gene Ontology}

The Gene Ontology provides a set of ontologies for molecular and
cellular biology, originally designed to support structured
annotations for genes and gene products in all species with respect to
molecular function (MF), biological process (BP) and cellular
component (CC). MF and BP both describe processes, but at different
spatiotemporal scales; in particular, BP includes processes that
unfold within cells and within tissues and organs of multicellular
organisms. Gene product annotations identify participants in the
processes.

Over time, GO development has increasingly emphasized a normalized
approach that includes supplementing existing human-readable text
description with formally specified explicit definitions for GO
classes. The formalization of GO is readily apparent in its
representation of biological regulation.

Regulatory processes may regulate other processes, at either the MF or
BP scale, or biological qualities. GO accordingly includes three broad
categories of regulation terms, regulation of molecular function,
regulation of biological process, and regulation of biological
quality. The first two are explicitly defined entirely with respect to
other GO terms, whereas the third comprises classes in which the
regulated qualities are specified by terms from PATO (see below) or
anatomy ontologies.

All GO regulation terms use one of three relations, regulates,
negatively_regulates and positively_regulates, to link regulation
terms to process or quality terms. The regulates relations are defined
in terms of qualities: a regulatory process causes a change in
magnitude to some quality, which in turn has an effect on the
frequency, rate or duration of some other type of process. Effects
that results in increases and decreases use positively_regulates and
negatively_regulates respectively. \cite{} (cite the paper on
cross-products in GO, PMID:20152934) The existing ontology structure
would also support the addition of subclasses to distinguish, for
example, regulation of the rate of a process from regulation of its
duration or time of onset.

quick intro...
